---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Text Analysis R Package

<!-- badges: start -->
<!-- badges: end -->

# Description

## Objective
Text analysis techniques such as Natural Language Processing (NLP) can swiftly 
extract relevant information from different corpus datasets such as documents, 
emails, SMS messages, social media and other textual resources. NLP encompasses 
some common techniques for analyzing human language such as tokenization, 
stemming and lemmatization, sentiment analysis and text classification.  

The goal of this R package "textAnalysisR" is to provide a series of functions 
that allows a user to analyze a text dataset using some common Natural Language 
Processing techniques, train some models and evaluate their performance.  

## Overview

In this package, we analyze the text dataset as follows:  

1. Load Data from text files
2. Install necessary packages  
3. Explore the text dataset  
4. Visualizing the corpus with word clouds
5. Data Preprocessing: Standardization and cleaning  
  - Convert to lower case   
  - Remove punctuations and numbers 
  - Remove extra white space  
  - Remove stop words
6. Data partitioning: Creating training and testing datasets  
7. Training a classifier on the data and evaluating model performance:  
  - Naïve Bayes 
  - Logistic Regression
  - Support Vector Machines (kernel= radial and linear)
  - Random Forest.  
6. Comparing results: Evaluate the model performance and compare them  


# Usage

## Programming Language

This package uses R as the main programming langauge and C++ for some scripts to
impove speed of the package.

## Installation

You can install the development version of textAnalysisR as follows:

``` {r}
# install.package("textAnalysisR")
```

Also, to get started on using this repository, type the following into your 
favorite git command line tool:

``` {r}
# git clone git@github.com:shaveen27/textAnalysisR.git
```

The previous command will download all of the code onto your computer. You will 
also need to download the data in study. You can find this in the *data* folder.


# Demo

There is a *tutorial* script to show how to use this package. We use the 
following dataset:

## Dataset

The SMS Spam Collection is a public set of SMS labeled messages that have been 
collected for mobile phone spam research.  

* SMS Spam Collection: 
  https://archive.ics.uci.edu/dataset/228/sms+spam+collection  

## Tutorial

This tutorial shows an example on how to use this package. It uses the dataset 
above and explore the different classifier mentioned in the description.
You cna find this tutorial under the name **textAnalysis_tutorial**.


### Information 

* The code in this repository was written by Aloka Dayarathne, Maha Moussa, 
Shaveen Britto, and Teresa White.

