---
title: "Short Introduction to text.analysis Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Short Introduction to text.analysis Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(tibble.print_min = 4L, tibble.print_max = 4L)
library(text.analysis)
set.seed(1234)
```

# Text Analysis Overview

## Introduction

Text analysis techniques such as Natural Language Processing (NLP) can swiftly 
extract relevant information from different corpus datasets such as documents, 
emails, SMS messages, social media and other textual resources. NLP encompasses 
some common techniques for analyzing human language such as tokenization, 
sentiment analysis and text classification.  

Working with text analysis using Natural Language Processing (NLP) involves 
several steps. Here's a general guide:

1. **Problem Definition:** Understand the objective of your text analysis. This 
could be to extract insights, classify text into categories, perform sentiment 
analysis, or something else.

2. **Data Collection:** Gather a corpus of text data relevant to your analysis. 
This could be from various sources such as websites, social media, books, 
articles, or any other text repositories.

3. **Data Preprocessing:** Implement data cleaning, normalization and 
standardization.
   
4. **Exploratory Data Analysis (EDA):** Understand the characteristics of your 
text data through statistical analysis, visualization, and summarization. This 
helps in identifying patterns and gaining insights.

5. **Model Training:** Split your data into training and testing sets. Train 
your chosen model(s) on the training data and tune hyperparameters to optimize 
performance.

6. **Model Selection:** Choose appropriate NLP models based on your problem and 
data characteristics. Common models include: Naive Bayes, Support Vector 
Machines, Random Forest, Logistic Regression.

7. **Model Evaluation:** Evaluate the performance of your trained model(s) using 
appropriate metrics for your task. For example, accuracy, precision, recall, 
F1-score, or Mean Squared Error (MSE) for regression tasks.


This document introduces you to the series of functions of the **text.analysis** 
package that allows a user to analyze a text dataset using some common Natural 
Language Processing techniques, train some models and evaluate their performance
.


## Description of Data

To explore the text analyis process of this package, we will use the dataset 
`SMSSpamCollection`. The SMS Spam Collection is a public set of SMS labeled 
messages that have been collected for mobile phone spam research. 

```{r setup}
library(text.analysis)

# use the dataset from the library
raw_data <- sms_spam_collection

raw_object <- convert_mail_list(raw_data)

# see description of the dataset
# ?sms_spam_collection
```

This raw dataset contains `r nrow(sms_spam_collection)` observations and 
`r ncol(sms_spam_collection)` columns: `category` and `message`. The variable 
`category` has two classes of messages: `ham` and `spam`. The variable `message`
shows a ham or spam text based on the `category` variable.

```{r}
# dimension of the dataset
dim(raw_data)

# example of some observation
head(raw_data)
```
For the raw data, there is no missing values and there are 4827 `ham` and 747
`spam` messages. In addition to this, the `min`, `mean` and `max` values for 
the **word count** in the messages for the observations are `1.00`, `15.18` and 
`171.00`, respectively. The `min`, `mean` and `max` values for the 
**message length** for the observations are `2.00`, `77.98` and `910`, 
respectively.

```{r}
# data exploration on the raw data
data_exploration <- explore_data(raw_object)
data_exploration[1:5]
```

## Data Preprocessing

The `message` feature in the raw dataset contains extraneous elements such as 
punctuation marks, whitespace characters, and numeric digits. These superfluous 
components can impede efficient analysis and modeling. Therefore, we want to 
remove these unnecessary elements.

In this phase of the analysis, we perform a comprehensive cleaning of the raw 
dataset by removing numeric values, punctuations, and stop words. We convert the 
text messages to lower case to implement data standardization. Also, we remove 
the white spaces to implement normalization.

```{r}
# Start cleaning
clean_corpus <- lower_case(raw_object)

clean_corpus <- remove_numbers(clean_corpus)

clean_corpus <- remove_punctuations(clean_corpus)

clean_corpus <- remove_whitespaces(clean_corpus)

clean_corpus <- remove_stopwords(clean_corpus)
```


Now we explore the clean dataset in a similar way we did for the raw data.

```{r}
# data exploration on the raw data
data_exploration <- explore_data(clean_corpus)
data_exploration[1:5]
```
We see that for the clean data, the `min`, `mean` and `max` values for 
the **word count** in the messages for the observations are `0.00`, `5.79` and 
`51.00`, respectively. The `min`, `mean` and `max` values for the 
**message length** for the observations are `0.00`, `35.05` and `290`, 
respectively. All the values are lower than for the raw data because of the 
cleaning implementation.


## Data Visualization

### Chart visualizatiom

Here are some visualization for the data exploration:

* The pie chart shows the total percentage of both classes: `ham` and `spam`. We 
see that 86.6% are classified as ham messages and 13.4% as spam messages.
* The bar charts show the 10 top more frequent words for the `ham` and `spam` 
messages. 
* For the `ham` messages, we have the following top words in a decreasing order: 
good, day, love, time, ü, lor, today, dont, send, pls. 
* For the `spam` messages, we have the following top words in a decreasing order
free, txt, mobile, text, claim, reply, prize, cash, nokia, send.


``` {r}
# explore clean data
explore_visuals(clean_corpus)
```


### Wordcloud visualization

Word clouds provide a visual representation of the most frequently occurring 
words or terms in a text corpus. This visualization tool will help us quickly 
identify the key words within the text data. It gives us a high-level overview 
of the prominent words. 

We generate the wordcloud visualization for the `ham` and `spam` sets.
  
``` {r}
# split the observations in ham, spam and all messages.
corpus_data <- split_spamham(clean_corpus)

# output: $Data, $Ham, $Spam
```

``` {r}
# wordcloud for the ham set
wordcloud_ham(corpus_data$Ham, min_freq = 50)
```

```{r}
# wordcloud for the spam set
wordcloud_spam(corpus_data$Spam, min_freq = 50)
```

By generating wordclouds for different subsets of the text data (ham and spam 
sets), we can compare and contrast the prominent words, potentially revealing 
insights into changing trends or preferences.

As we can see, the 10 top words for both sets appear on these wordcloud 
visualizations as well.


## Splitting Data

We split data into training, and test sets for several important reasons for 
modelling. One of the reasons is to avoid overfitting. By separating the data 
used to train the model (training set) from the data used to evaluate its 
performance (test sets), we can assess how well the model generalizes to unseen 
data. This helps prevent overfitting, where the model learns the noise and 
peculiarities of the training data too closely, leading to poor performance on 
new data. We also use some default hyperparameter Tuning to select the best 
performing model to avoid overfitting and ensure the model's generalization 
capability.

We create the model matrix to use modelling process. In this step, the resulting 
output of the function is a data matrix where the initial column denotes the 
message type, while the subsequent columns represent the words in the dataset.
Then, we partition our clean data as follows: 70% training and 30% test sets. 
This split creates the training and test data matrices. 

``` {r}
# model matrix
final_model_corpus <- final_model_df(clean_corpus, tol = 10)

set.seed(1234)
# partition
split_data <- partition(model_df = final_model_corpus, prob = 0.7)
```


Here is a table with the number of observations for the training and test sets:

```{r}
# Print the table
split_table <- matrix(c(3379, 523, 3902, 1448, 224, 1672, 4827, 747, 5574),
  nrow = 3, ncol = 3
)

# Assign row and column names
row_names <- c("Ham", "Spam", "Total")
col_names <- c("Training", "Test", "Dataset")
dimnames(split_table) <- list(row_names, col_names)

# Print the matrix
split_table
```

## Classification Models


We applied four classifiers to our data: Naive Bayes, Support vector machines 
(SVMs), Random Forest, Logistic regression.

### Naive Bayes Classification

Naive Bayes is a straightforward yet potent classification algorithm widely 
utilized in machine learning and data analysis. Its main principles hinge on two 
key ideas: conditional independence and Bayes' Theorem. The model assumes that 
features are unrelated to each other given the class variable, simplifying 
computations. It then calculates the likelihood of a class based on input 
features, making predictions by selecting the class with the highest likelihood. 
Despite its basic assumptions, Naive Bayes often proves remarkably effective, 
especially for complex datasets. Its efficiency and quick training make it a 
practical choice for large-scale applications.

### Support vector machines Classification (SVMs)

Support vector machines (SVMs) stand out as a potent tool in supervised machine 
learning, serving for classification and regression tasks alike. Their core aim 
is to pinpoint an optimal hyperplane that maximizes the margin between distinct 
classes of data points. Operating by projecting input data into a 
higher-dimensional feature space, SVMs locate a linear decision boundary with 
the aid of kernel functions, adeptly handling non-linear classification 
quandaries. Among these functions are the linear, polynomial, and radial basis 
function (RBF) kernels, each selected based on data characteristics and decision 
boundary complexity. Noteworthy for their prowess in high-dimensional realms, 
memory thriftiness by utilizing a subset of training points, and versatility 
across diverse domains such as text classification, image recognition, and 
bioinformatics, SVMs remain a stalwart choice in the machine learning toolkit.

### Random Forest Classification

Random Forest is a strong method in machine learning that combines many decision 
trees to make better predictions. It works like this: first, it builds a 
"forest" of decision trees. Each tree is trained on a random part of the data 
and only looks at some of the features. For deciding the class in classification 
tasks, it takes the most common choice among the trees. In regression, it 
averages the predictions of all trees. This randomness helps to avoid a common 
problem called overfitting, where the model is too complex. Random Forest can 
handle different types of data, like numbers, categories, and text. It also 
tells us which features are most important for making predictions. Another good 
thing about Random Forest is that it performs well in many situations without 
needing a lot of adjustments. Overall, Random Forest is a reliable method in 
machine learning because it combines many simple trees to give accurate and 
understandable predictions, making it popular for many tasks.

### Logistic regression Classification

Logistic regression is a useful statistical method for solving binary 
classification tasks, like predicting whether a customer will buy a product or 
if a transaction is fraudulent. It works by modeling the probability of the 
binary outcome based on predictor variables, using a sigmoid function to 
transform them into probabilities between 0 and 1. The model parameters, which 
are coefficients, are estimated through maximum likelihood estimation, finding 
values that best match the observed data. These coefficients represent changes 
in the log odds of the outcome for each unit change in the predictor, and their 
exponentiated values are odds ratios, easier to understand. To check how well 
the model fits the data, tests like the Hosmer-Lemeshow test are used, along 
with metrics like accuracy and recall. Logistic regression can sometimes overfit
, especially with many predictors, but techniques like L1 or L2 regularization 
can help prevent this. Overall, logistic regression is a popular and 
understandable method for binary classification, providing insight into the 
relationship between predictors and outcomes.

**Overall**

Because of their special qualities and capacity to deal with the difficulties 
presented by text data such as its high dimensionality, sparse characteristics, 
and intricate word relationships these algorithms are frequently employed for 
text categorization. The algorithm of choice is frequently determined by the 
particulars of the text data as well as the needs of the application, including 
interpretability, computational effectiveness, and performance on challenging 
tasks.


## Results

In this part of the analysis, we will explore four classifiers on the data. For 
all of them, we compute the confusion matrix and accuracy measures.

A confusion matrix is a powerful tool for evaluating the performance of any 
classification model, whether it's a binary or multi-class problem. It provides 
a detailed breakdown of the model's predictions compared to the actual ground 
truth labels. The confusion matrix is a table with rows representing the true 
classes and columns representing the predicted classes. The diagonal elements 
represent the correctly classified instances, while the off-diagonal elements 
represent the misclassified instances.

For the accuracy measures, various performance metrics can be calculated:

+ True Positives (TP): The number of instances correctly predicted as positive.
+ False Positives (FP): The number of instances incorrectly predicted as positive.
+ True Negatives (TN): The number of instances correctly predicted as negative.
+ False Negatives (FN): The number of instances incorrectly predicted as negative.

From these values, various performance metrics can be calculated:

**Calculations**
$$Accuracy = \frac{TP + TN}{TP + FP + FN + TN}$$
$$Precision = \frac{TP}{TP + FP}$$
$$Recall \ (Sensitivity) = \frac{TP}{TP + FN}$$
The formula for calculating the F1 score is:

$$F1-Score = \frac{2 * precision * recall}{precision + recall}$$


### Naïve Bayes classification  

We run the model and get the confusion matrix and the accuracy measures 
mentioned.

``` {r}
# for example: Naïve Bayes classifier
nb_model <- nb_classification(split_data)

# output for Naïve Bayes: Confusion Matrix
nb_model$Confusion_Matrix
```

``` {r}
# output for Naïve Bayes: Accuracy_Measures
nb_model$Accuracy_Measures
```


### Support Vector Machine classification 

We run the model and get the confusion matrix and the accuracy measures 
mentioned.

``` {r}
# for example: SVM classifier
svm_model <- svm_classification(split_data, kernel = "linear")

# output for SVM: Confusion Matrix
svm_model$Confusion_Matrix
```
``` {r}
# output for SVM: Accuracy_Measures
svm_model$Accuracy_Measures
```


### Random Forest classification model. 

We run the model and get the confusion matrix and the accuracy measures 
mentioned.

```{r}
# for example: RF classifier
rf_model <- rf_classification(split_data)

# output for RF: Confusion Matrix
rf_model$Confusion_Matrix
```

``` {r}
# output for RF: Accuracy_Measures
rf_model$Accuracy_Measures
```


### Logistic Regression classification model. 

We run the model and get the confusion matrix and the accuracy measures 
mentioned.

```{r}
# for example: LOG classifier
log_model <- log_classification(split_data)

# output for LOG: Confusion Matrix
log_model$Confusion_Matrix
```


``` {r}
# output for Naïve Bayes: Accuracy_Measures
log_model$Accuracy_Measures
```


## Comparison of Model

Here we compare the model performance of the four classifiers:

```{r}
cbind(
  Model = c(
    "Naive Bayes", "Linear - SVM",
    "RandomForest", "Logistic Regression"
  ),
  rbind(
    nb_model$Accuracy_Measures, svm_model$Accuracy_Measures,
    rf_model$Accuracy_Measures, log_model$Accuracy_Measures
  )
)
```

When comparing the performance of the classification models applied to the SMS 
dataset, distinct patterns emerge regarding their efficacy in distinguishing 
between `ham` and `spam` messages. **Naive Bayes** showcases a streamlined 
approach, leveraging conditional independence and Bayesian principles to achieve 
high accuracy. **Support Vector Machines**, particularly the linear variant, 
excel in delineating linear decision boundaries, making them adept at handling 
high-dimensional text data. **Random Forest**, with its ensemble of decision 
trees, provides robustness against overfitting and performs reliably across 
various types of data. In contrast, **Logistic Regression** offers 
interpretability but may struggle with overfitting, especially in scenarios with
numerous predictors. Each model brings its own strengths and considerations to 
the table, highlighting the importance of selecting the most suitable approach 
based on the specific requirements and constraints of the task at hand.



## Discussion & Conclusions

The analysis presented provides a comprehensive evaluation of four 
classification models applied to the task of distinguishing between `ham` and 
`spam` messages within the SMS dataset. Each model offers unique strengths and 
considerations, influencing their performance across various metrics such as 
accuracy, precision, recall, and F1-score.

Naive Bayes, a simplistic yet powerful algorithm, demonstrates notable 
performance with high accuracy (98.24%), precision (0.97), recall (0.90), and 
F1-score (0.93). Its reliance on conditional independence and Bayes' Theorem 
makes it particularly efficient for large-scale applications, making it a 
pragmatic choice for text categorization tasks. The model's ability to swiftly 
process data and make predictions based on probabilistic principles underscores 
its effectiveness in classifying ham and spam messages.

Support Vector Machines (SVMs), specifically the linear SVM variant used in this analysis, also emerge as a formidable contender, boasting impressive accuracy 
(97.75%), precision (0.95), recall (0.89), and F1-score (0.92). Linear SVMs 
excel in delineating linear decision boundaries, making them particularly 
suitable for high-dimensional datasets like text data. Their efficiency in 
processing large volumes of data and their ability to handle sparse features 
contribute to their robust performance in text classification tasks.

Random Forest, an ensemble method that combines multiple decision trees, 
delivers commendable results with an accuracy of 97.21% but slightly lower 
precision (0.93), recall (0.86), and F1-score (0.89) compared to Naive Bayes and 
SVMs. Random Forest's strength lies in its ability to mitigate overfitting by 
aggregating predictions from multiple trees, thereby enhancing model robustness 
and generalization. Its capacity to handle different types of data and identify 
important features makes it a reliable choice for a wide range of classification 
tasks.

In contrast, Logistic Regression, while exhibiting respectable accuracy (93.33%)
, lags behind in precision (0.71), recall (0.86), and F1-score (0.78) compared 
to the other models. Despite its simplicity and interpretability, Logistic 
Regression may struggle with overfitting, particularly in scenarios with many 
predictors. However, techniques such as L1 or L2 regularization can mitigate 
this issue, enhancing the model's performance and generalization capability.

Ultimately, the selection of the most appropriate classification model hinges on 
various factors such as interpretability, computational efficiency, and 
performance on challenging tasks. Naive Bayes and linear SVMs stand out for 
their commendable performance in text categorization tasks, while Random Forest 
offers robustness against overfitting, and Logistic Regression provides 
valuable insights into binary classification tasks. Therefore, the choice of 
model should be tailored to the specific requirements and constraints of the 
application at hand, ensuring optimal performance and effectiveness in 
addressing the task objectives.
